---
title: "Practicing Basic Analytical Techniques on a Small Dataset using R"
author: "Carrington Body"
output: word_document
---

&nbsp;

<font size="4">
**Synopsis:** *This is a small project consisting of exercises that will allow me to practice elementary techniques that you can use in R to analyze data sets. These exercises range from making a data frame from reading in a data set to manipulating the data frame and making calculations from specific data.*

&nbsp;

**Creating a Uniform Histogram** Here, I just want to see if I can generate a surplus of random numbers (roughly 100,000) from a uniform distribution from 5 to 20. This means that every random number generated will be a number in the range of 5 to 20, shown on a histogram with a roughly uniform distribution. I will not use ggplot for this exercise.

```{r}
set.seed(555)
unifrandvec <- runif(n = 100000, min = 5, max = 20)
hist(unifrandvec, col = "red")
```

&nbsp;

**EPA Vehicle Info Data set** For these exercises, I will use the 2021 fuel 
economy data set from the EPA (Environmental Protection Agency). More details are available at the following link:
https://www.openintro.org/data/index.php?data=epa2021


**Reading in the file** Here, I will read in the csv file and assign the output to a data frame for us to view in the R workspace. I also want to make the variables of the data frame directly accessible in the workspace so that I can manipulate them with different R techniques.

```{r}
epa2021 <- read.csv("epa2021.csv", header = TRUE)
#attaching file to access variables
attach(epa2021)
#checking structure of dataframe
str(epa2021)
#viewing dataframe
View(epa2021)
```

&nbsp;

**Observations Based on Different Conditions** In this exercise, I will calculate the fraction of cars in this data set that satisfy the following conditions:

* 1) that have a city mpg of 30 or more.
```{r}
total_cars <- dim(epa2021)[1]
city30 <- sum(epa2021$city_mpg >=30) / total_cars
city30
#Per 100 cars, at least 6 of them will have a city mpg of 30 miles or more.
```

* 2) whose highway mpg is less than their city mpg.
```{r}
sum(epa2021$hwy_mpg < epa2021$city_mpg) / total_cars
#Per 100 cars, about 2 of them will have a highway mpg that is less than their city mpg.
```
* 3) that have engine displacement that is at least 3.5 or whose highway mpg is 
strictly less than 30.
```{r}
OrSum <- sum(epa2021$engine_displacement >= 3.5 | epa2021$hwy_mpg < 30) / total_cars
OrSum
#Per 100 cars, there would be about 68 cars that have an engine displacement that is at least 3.5 or a highway mpg that is less than 30 miles.
```

&nbsp;

**Mean, SD, & Quantiles of Porsches** Here, I am getting the mean, standard deviation and the quantiles of the city mpg of all Porsche cars. This will help me analyze specific data that pertain to singular types of entities.

```{r}
porsches <- subset(epa2021, mfr_name=="Porsche")
View(porsches)
#####################
paste("Mean of city mpg for all porsches:", mean(porsches$city_mpg))
paste("Standard Deviation of city mpg for all porsches:",sd(porsches$city_mpg))
quantile(porsches$city_mpg)
```

&nbsp;

**More Fractions** Here, I will determine the fraction of the Porsche cars that 
have a city mpg of at least 20.

```{r}
total_porsche <- dim(porsches)[1]
porsche.citympg.atleast20 <- sum(porsches$city_mpg >= 20) / total_porsche
paste("Fraction of Porsches that have a city mpg of at least 20 miles:", porsche.citympg.atleast20)
print("This means that per 100 Porsches, there will be at least 12 of them that have a city mpg of at least 20 miles.")
#0.126984126984127 * 63
print("Since there are only 63 Porsches in this data set, only 8 of them have a city mpg of at least 20 miles.")
```

&nbsp;

**Smaller Data Frame of Porsches & Ferraris** I will create a smaller data frame consisting 
only of the Porsche and Ferrari cars in this data set and 
whose variables consist of mfr_name, engine_displacement, no_cylinders,
city_mpg, and highway_mpg. This will also help me analyze different levels of data with a selection of variables within a data set.

```{r}
porsches <- subset(epa2021, mfr_name=="Porsche")
ferraris <- subset(epa2021, mfr_name=="Ferrari")
PorFer <- rbind(porsches,ferraris)
PorFer[c('mfr_name', 'engine_displacement', 'no_cylinders', 'city_mpg', 'hwy_mpg')]
```

&nbsp;

**Sorting Data Frame in Decreasing Order of City Mpg**

```{r}
rev(order(PorFer$city_mpg)) #Displays the indexes of car entries in decreasing order
PorFer[rev(order(PorFer$city_mpg)),]
```

&nbsp;

**Combined Averages of City mpg & Highway mpg** The EPA rating for combined mpg presumes that we drive 55 percent of the time in the city and 45 percent of the time on the 
highway, so I am calculating a weighted (combined) average of the city mpg and 
highway mpg using weights of 0.55 and 0.45, respectively, for *only Ferraris* in this 
data set.

```{r}
#0.55 x Average of city mpg + 0.45 x Average of hwy mpg
ferraris <- subset(epa2021, mfr_name=="Ferrari")
combined.avg <- round(0.55 * mean(ferraris$city_mpg) + 0.45 * mean(ferraris$hwy_mpg)) #rounding to the nearest whole number
paste("Combined weighted average of city mpg and highway mpg of all Ferraris in this data set:", combined.avg, "miles")
```

&nbsp;

**Creating a Bar Plot for Drive System variable (drive_sys) w/ its Respective Counts**

```{r}
tab <- table(epa2021["drive_sys"])
xcenter <- barplot(tab,col="green",ylim = c(0,400),xlab="Drive System",ylab="Count")
counts <- as.integer(tab)
vertadjust <- 2
text(xcenter,counts+vertadjust,labels=as.character(counts))
```

&nbsp;

**Side-by-side Boxplots** Here, I am creating side-by-side boxplots comparing the 
highway mpg of vehicles that are guzzlers and those that are not.

```{r}
boxplot(epa2021$hwy_mpg~epa2021$guzzler, col = "yellow", xlab = "Is a guzzler?", ylab = "Highway mpg")
legend(2,60,legend = c("n = no", "y = yes"))
```
</font> 